% !TEX TS-program = pdflatex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[sn-mathphys,pdflatex,iicol]{sn-jnl}% Math and Physical Sciences Reference Style

%%%% Standard Packages
%\usepackage{polyglossia} % Must come before biblatex

\usepackage{bm}
\usepackage{csquotes}
%\usepackage{fontspec}
\usepackage{hyperref}
%\usepackage{lua-visual-debug}
\usepackage{tabularx}
%%%%

\jyear{2022}

%% as per the requirement new theorem styles can be included as shown below
%%\theoremstyle{thmstyleone}%
%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
%%\newtheorem{proposition}[theorem]{Proposition}%
%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.

%%\theoremstyle{thmstyletwo}%
%%\newtheorem{example}{Example}%
%%\newtheorem{remark}{Remark}%

%%\theoremstyle{thmstylethree}%
%%\newtheorem{definition}{Definition}%

\raggedbottom
%%\unnumbered% uncomment this for unnumbered level heads

%\setdefaultlanguage{english}
%\setotherlanguage{czech}

\hypersetup{
	pdfencoding=auto,
	unicode=true,
	bookmarksopen=true,
	bookmarksopenlevel=3
}

\newcommand{\name}[1]{\textit{#1}}
\newcommand{\mathfield}{\ensuremath{\mathbb}}
\newcommand{\mathmat}{\ensuremath{\mathbf}}
\newcommand{\mathset}{\ensuremath{\mathbb}}
\newcommand{\mathspace}{\ensuremath{\mathcal}}
\newcommand{\mathvec}{\ensuremath{\bm}}

\newcounter{enumroman}
\newenvironment{romanitems}{\begin{list}{\bfseries(\roman{enumroman})\hfill}{\usecounter{enumroman}\setlength{\labelwidth}{\leftmargin}\addtolength{\labelwidth}{-1\labelsep}\topsep=0mm plus 2pt\itemsep=0mm\parsep=0mm plus 2pt\itemindent=0mm}}{\end{list}}

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}

\begin{document}

\title[Adaptive graph coarsening]{Balancing performance and complexity with adaptive graph coarsening}

\author*[1,2]{\fnm{Marek} \sur{Dědič}}\email{marek@dedic.eu}

\author[2]{\fnm{Lukáš} \sur{Bajer}}\email{lubajer@cisco.com}

\author[2]{\fnm{Pavel} \sur{Procházka}}\email{paprocha@cisco.com}
\author[3]{\fnm{Martin} \sur{Holeňa}}\email{martin@cs.cas.cz}

\affil[1]{\orgdiv{Faculty of Nuclear Sciences and Physical Engineering}, \orgname{Czech Technical University in Prague}, \orgaddress{\street{Břehová 7}, \city{Prague}, \postcode{110 00}, \country{Czech Republic}}}

\affil[2]{\orgdiv{Cognitive Intelligence}, \orgname{Cisco Systems, Inc.}, \orgaddress{\street{Karlovo náměstí 10}, \city{Prague}, \postcode{120 00}, \country{Czech Republic}}}

\affil[3]{\orgdiv{Institute of Computer Science}, \orgname{Czech Academy of Sciences}, \orgaddress{\street{Pod vodárenskou věží 2}, \city{Prague}, \postcode{182 07}, \country{Czech Republic}}}

\input{src/abstract.tex}

\maketitle

\input{src/introduction.tex}
\input{src/related-work.tex}
\input{src/performance-complexity.tex}
\input{src/harp-framework.tex}
\input{src/our-method.tex}
\input{src/experimental-evaluation.tex}
\input{src/conclusion.tex}

\section*{Declarations}

\subsection*{Funding}

The research reported in this paper has been supported by the German Research Foundation (DFG) funded project 467401796.

Marek Dědič, Lukáš Bajer and Pavel Procházka are employed by Cisco Systems, Inc. Martin Holeňa is employed by the Institute of Computer Science of the Czech Academy of Sciences, the Czech Technical University in Prague and the University of Rostock.

\subsection*{Conflicts of interest/Competing interests}

Martin Holeňa is a reviewer for the 2023 ECML journal track.

\subsection*{Ethics approval}

Not applicable.

\subsection*{Consent to participate}

Not applicable.

\subsection*{Consent for publication}

Not applicable.

\subsection*{Availability of data and material}

All used datasets are publicly available.


\subsection*{Authors' contributions}

All authors contributed to the research conception and design. Experiment programming and execution was done by Marek Dědič. Statistical interpretation of the results was done by Marek Dědič and Martin Holeňa. The first draft of the manuscript was written by Marek Dědič and all authors commented on previous versions of the manuscript. All authors read and approved the final manuscript.

\bibliography{zotero}

\end{document}
